---
title: "Uncertain emotion discrimination differences between musicians and nonmusicians is determined by fine structure association: Hilbert transform psychophysics"
date: "Updated on 15 August, 2019"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Authors:** Francis A.M. Manno, Raul R. Cruces, Condon Lau, Fernando A. Barrios   
**Contact:** francis.manno@sydney.edu.au, raulrcruces@inb.unam.mx, fbarrios@unam.mx, condon.lau@cityu.edu.hk  
  
**OSF:** https://osf.io/8ws7a  
**Keywords:** emotion; psychophysics; modulation; fine structure; envelope; frequency; amplitude  
  
# Abstract
Humans perceive musical sound as a complex phenomenon, which is known to induce an emotional response. The cues used to perceive emotion in music have not been unequivocally elucidated. Here, we sought to identify the attributes of sound that confer an emotion to music and determine if professional musicians have different musical emotion perception than non-musicians. The objective was to determine which sound cues are used to resolve emotional signals. Happy or sad classical music excerpts modified in fine structure or envelope conveying different degrees of emotional certainty were presented. Certainty was determined by identification of the emotional characteristic presented during a forced-choice discrimination task. Participants were categorized as good or poor performers (n = 32, age 21.16 ± 2.59 SD) and in a separate group as musicians in the first or last year of music education at a conservatory (n = 32, age 21.97 ± 2.42). We found that temporal fine structure information is essential for correct emotional identification. Non-musicians used less fine structure information to discriminate emotion in music compared with musicians. The present psychophysical experiments revealed what cues are used to resolve emotional signals and how they differ between non-musicians and musically educated individuals.


# Code and Methods  
The code for **figure 2** and **figure 4** can be found in `Final_figures.R`. 

## R 3.4.4, required packages
```{r packages, echo=TRUE,message=FALSE, warning=FALSE}
library(gplots)
library(gridExtra)
library(MASS)
library(pander)
library(magrittr)
library(dplyr)
library(ggplot2)
library(RCurl)
library(scatterplot3d)
library(klaR)
library(pander)
library(candisc)
```

# Group Heatmaps
This was created calculating the accuracy of each subject per stimuli. First, stimuli where binarized based in the answer of **Fine structure emotion**, with 1 if is correct and 0 when wrong. The total was divided by the total amount of stimuli or each category (nb0...nb64).  

## Accuracy for both emotions: All subjects
```{r Heatmap Accuracy, echo=FALSE, warning=FALSE, message=FALSE}
source('http://www.sthda.com/sthda/RDoc/functions/addgrids3d.r')
# plot grid on a 3d scatterplot
# Get functions from source: github.com/rcruces/R-plots/master/R-colors/colormaps_functions.R
script <- getURL("https://raw.githubusercontent.com/rcruces/R-plots/master/R-colors/colormaps_functions.R",ssl.verifypeer = FALSE)
eval(parse(text = script)); rm(script)

# Set path, change this to your home/githubDirectory path
home="~/"
setwd(paste0(home,"git_here/2019_uncertain_emotion_discrimination"))

# Read the Subjects Information
cases <- read.csv("database/cases.csv")

# Reads the subjects Responses
result <- read.csv("database/cases_results.csv")

# Slices Happy and Sad
hap <- result[result$emo=="h",]
sad <- result[result$emo=="s",] 

# Creates a matrix with the Percentage of Correct Responses per FINE STRUCTURE stimuli
calAccu <- function(df){
  accuracy <- c()
  for (i in unique(df$fts)) {
    sub.cases <- df[df$fts==i,]
    X <- apply(sub.cases[,7:70],2,function(x) sum(ifelse(sub.cases$emo.fts == x, 1,0))/length(x) )
    accuracy <- cbind(accuracy,X)
  }
  colnames(accuracy) <- unique(df$fts)
  return(as.matrix(accuracy))  
}
# Color-palette
colG<-c("navy","dodgerblue4","dodgerblue2","deepskyblue","lightskyblue","red","firebrick2","firebrick3","firebrick","firebrick4")
# Color-map
col.map <- optim.color(calAccu(result),colG)
# Heatmap of all subjects, Clustered by Group
heatmap.2(calAccu(result),Colv=FALSE,main= "Fine Structure Accuracy",
          density.info="histogram",
          dendrogram='row',
          trace="none",
          tracecol="white",
          key.title = "Discriminability",     
          col    = col.map$color.palette,   # Colormap
          breaks = col.map$palette.breaks,
          srtCol=1,  
          RowSideColors = c("mediumpurple4","darkolivegreen","darkolivegreen3","mediumpurple")[cases$class])
```
  
\newpage

## Table 1. 
```{r tabla, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4, fig.height= 4}
cases$accu <- calAccu(result)[,1]
cases$subclass <- paste0(cases$gender,".",cases$class)

tabla <- cbind(table(cases$subclass),
               round(aggregate(age~subclass, data=cases, mean)[,2],2), round(aggregate(age~subclass, data=cases, sd)[,2],2),
               round(aggregate(music.yrs~subclass, data=cases, mean)[,2],2), round(aggregate(music.yrs~subclass, data=cases, sd)[,2],2),
               round(aggregate(age~subclass, data=cases, range)[,2],1),
               round(aggregate(accu~subclass, data=cases, mean)[,2]*100,2), round(aggregate(accu~subclass, data=cases, sd)[,2]*100,2),
               round(aggregate(accu~subclass, data=cases, mean)[,2]*32,2)
               )

colnames(tabla) <- c("N","age.mean", "age.sd", "years.mean", "years.sd", "range.1","range.2", "correct.mean","correct.sd","correct.N")

tabla <- tabla[c(5,1,8,4,7,3,6,2),]
rownames(tabla) <- c("M.good","F.good","M.poor","F.poor","M.first","F.first","M.last","F.last")

pander(tabla)
```
  
  
## Fine Structure Accuracy by Gender  
```{r heatmap.gender, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4, fig.height= 4}
# FUNCTION for heatmap plotting
hmap <- function(Mtx, Title){
heatmap.2(Mtx,Colv=FALSE,main= Title,
            density.info="histogram",Rowv = TRUE,
            dendrogram='row',
            trace="row",
            tracecol=NA,
            key.title = "Discriminability",     
            col    = col.map$color.palette,   # Colormap
            breaks = col.map$palette.breaks,
            srtCol=1)}

library(corrplot)
indx <- which(cases$gender=="M")
M.sad <- calAccu(hap)[indx,2:7]
M.hap <- calAccu(sad)[indx,2:7]
indx <- which(cases$gender=="F")
F.sad <- calAccu(hap)[indx,2:7]
F.hap <- calAccu(sad)[indx,2:7]
# Heatmap: Happy-Sad
hmap(M.sad,"Males: Sad")
hmap(F.sad,"Females: Sad")
hmap(M.hap,"Males: Happy")
hmap(F.hap,"Females: Happy")
```

# Figure 2. Accuracy and discriminability concerning Happy and Sad   
```{r fig2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4}
acu.sad <- cbind(calAccu(sad)[,1],calAccu(hap)[,2:7])
acu.hap <- cbind(calAccu(hap)[,1],calAccu(sad)[,2:7])
heatmap.2(acu.hap,Colv=FALSE,Rowv = FALSE,main= "Happy Fine structure",
          density.info="histogram",
          dendrogram='none',
          trace="none",
          tracecol="white",
          key.title = "Discriminability",     
          col    = col.map$color.palette,   # Colormap
          breaks = col.map$palette.breaks,
          srtCol=1,  
          RowSideColors = c("mediumpurple4","darkolivegreen","darkolivegreen3","mediumpurple")[cases$class])

heatmap.2(acu.sad,Colv=FALSE,Rowv = FALSE,main= "Sad Fine Structure",
          density.info="histogram",
          dendrogram='none',
          trace="none",
          tracecol="white",
          key.title = "Discriminability",     
          col    = col.map$color.palette,   # Colormap
          breaks = col.map$palette.breaks,
          srtCol=1,  
          RowSideColors = c("mediumpurple4","darkolivegreen","darkolivegreen3","mediumpurple")[cases$class])

```
  
  
```{r fig2.hist, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2.5}
par(mfrow=c(1,2))
hist(acu.hap,ylim=c(0,160),breaks = 10,main="Happy Discriminability", col=col.map$color.palette[c(1,1,1,2,3,4,5,6,7,7)],border = "white")
hist(acu.sad,ylim=c(0,160),breaks = 10,main="Sad Discriminability", col = col.map$color.palette[c(1,1,1,2,3,4,5,6,7,7)],border = "white")

```
  
## Accuracy calculaded by subjects and group  
  
```{r Lines-dots, echo=FALSE, warning=FALSE, message=FALSE}
accuracy <- calAccu(result)
blank.plot <- function (X,lab,Main) {
  Ax.Col="black"
  Main <- as.character(Main)
  plot(X,ylim=c(0,1),xlim = c(0,length(X)+1),pch=19,ylab="score",main=Main,col=NA,xaxt='n',bty='n',xlab="",axes = FALSE,
       las=1,col.axis=Ax.Col,col.lab=Ax.Col,col.main=Ax.Col,cex.axis=1.2,cex=seq(0.5,1.5,length.out = 35),cex.main=2)
  polygon(c(0,0,8,8,0),c(0,1,1,0,0),col="gray85",border = NA)
  abline(h=seq(0,1,0.2),lty=1,col="white")
  axis(1, at=1:7, col.axis=Ax.Col,labels=lab, lty=1, col=Ax.Col, las=1,lwd=3,cex.axis=1,col.lab=Ax.Col)
  axis(2, at=seq(0,1,0.2), labels=seq(0,1,0.2), lty=1, col.axis=Ax.Col,las=1,lwd=3,cex.axis=1.2,col.lab=Ax.Col,col=Ax.Col)
}
lines.plot <- function(y,Col,Tilte) {
  x <- apply(y,2,mean)
  blank.plot(x,lab=colnames(accuracy),Main = Tilte)
  apply(y,1, function(x) lines(x,col=Col,lwd=0.5))
  lines(x,col=Col,lwd=2.5)
  points(x,bg=Col,lwd=2,cex=1.5,pch=21,col="white")
  
}
par(mfrow=c(2,2))
# FEMALE PERFORMANCE
lines.plot(accuracy[cases$class == "poor" & cases$gender=="M",],"mediumpurple", "Poor Male")
lines.plot(accuracy[cases$class == "poor" & cases$gender=="F",],"mediumpurple", "Poor Female")
# MALE PERFORMANCE
lines.plot(accuracy[cases$class == "best" & cases$gender=="M",],"mediumpurple4", "Good Male")
lines.plot(accuracy[cases$class == "best" & cases$gender=="F",],"mediumpurple4", "Good Female")
# HIGH PERFORMANCE
lines.plot(accuracy[cases$class == "high" & cases$gender=="M",],"darkolivegreen", "High Male")
lines.plot(accuracy[cases$class == "high" & cases$gender=="F",],"darkolivegreen", "High Female")
# LOW PERFORMANCE
lines.plot(accuracy[cases$class == "low" & cases$gender=="M",],"darkolivegreen3", "Low Male")
lines.plot(accuracy[cases$class == "low" & cases$gender=="F",],"darkolivegreen3", "Low Female")
```
  
\clearpage
  
## ANOVA: Comparisons between groups  
Is accuracy in the response given by the class belonging and gender?  
```{r anova, echo=FALSE,message=FALSE, warning=FALSE}
require(pander)
mod1 <- aov(accuracy[,1]~factor(cases$class)+factor(cases$gender))
pander(summary(mod1),caption = "ANOVA, accuracy by class")
pander(TukeyHSD(mod1),caption = "ANOVA post hoc, accuracy by class")
```
  
\clearpage
  
# Linear Discriminant Analysis: Happy  
 
$Group\~nb0+nb2+nb4+nb8+nb16+nb32+nb64$  

```{r lda, echo=FALSE,warning=FALSE,message=FALSE}
# Data Reshape
d.happy <- data.frame(cbind(cases$class,cases$age,cases$music.yrs,cases$gender,acu.hap)) 
colnames(d.happy) <- c("type","age","music.yrs","gender","nb0","nb2","nb4","nb8","nb16","nb32","nb64")
d.happy$type <- cases$class
d.happy$id <- rownames(d.happy)

# Linear Discriminant model MASS library
m.happy <- lda(type~nb0+nb2+nb4+nb8+nb16+nb32+nb64, d.happy)
# Obtains the predictor from the LDA to an object
p.happy <- predict(m.happy)
# Accuracy of prediction
freqtable <- table(p.happy$class, d.happy$type)
# Percent correct of each category of G
pander(diag(prop.table(freqtable)),"Accuracy prediction: Happy")
# Frecuency table
pander("  \n  ")
rownames(freqtable) <- paste0("Predicted ", d.happy$type %>% levels)
freqtable %>% addmargins %>% pander("LDA happy: Observed vs. Predicted Frequencies")
pander("  \n  ")

# Total percent correct
pander(paste("Happy total % correct:",sum(diag(prop.table(freqtable)))))
pander("  \n  ")

# Proportions
prop.table(freqtable) %>% addmargins %>% pander("LDA sad: Observed vs. Predicted Proportions %")
pander("  \n  ")

# LDA to variables
d.happy$LDA1 <- p.happy$x[,1]
d.happy$LDA2 <- p.happy$x[,2]
d.happy$LDA3 <- p.happy$x[,3]

# Table
gMeans <- d.happy %>% group_by(type) %>% select(LDA1,LDA2,LDA3) %>%  summarise_each(funs(mean)) 
gMeans %>% pander("Happy: group means by LDA")
pander("  \n  ")

# Plot the LDA
grid.arrange(
ggplot(d.happy, aes(LDA1, fill = type)) + geom_density(alpha = 0.5, color = NA) + theme(legend.position = "top"),
ggplot(d.happy, aes(LDA2, fill = type)) + geom_density(alpha = 0.5, color = NA) + theme(legend.position = "top"),
ggplot(d.happy, aes(LDA3, fill = type)) + geom_density(alpha = 0.5, color = NA) + theme(legend.position = "top"),
# Plot the LDA1 vs LDA2 with color per group
ggplot(d.happy, aes(LDA1, LDA2, color = type)) + 
  geom_point(alpha = 0.5) + 
  geom_text(data = gMeans, 
            aes(label = type),
            color = "black", 
            vjust = 1.75) + 
  geom_point(data = gMeans, 
             aes(fill = type), 
             size = 4, 
             color = "black", 
             pch = 21) + 
  theme(legend.position = "none") +
  coord_equal(),
ncol=2, nrow = 2)
```

  
```{r lda.happy, echo=FALSE,warning=FALSE,message=FALSE, fig.height=2.5}
# Plots the discriminat functions
par(mfrow=c(1,3))
# scater plot of each subject by LDA value
scatterplot3d(d.happy$LDA1,d.happy$LDA2,d.happy$LDA3,color = c("red","green","blue","orange")[d.happy$type],pch = 19,box = FALSE,main = "LDA Happy", xlab = "LDA1", ylab = "LDA2", zlab = "LDA3")
addgrids3d(d.happy$LDA1,d.happy$LDA2,d.happy$LDA3, grid = c("xy", "xz", "yz"))
# Linear discrimant functions
drawparti(grouping = d.happy$type, x = d.happy$LDA1, y = d.happy$LDA2, xlab = "LDA1", ylab = "LDA2")
drawparti(grouping = d.happy$type, x = d.happy$LDA2, y = d.happy$LDA3, xlab = "LDA2", ylab = "LDA3")
```
  
\clearpage
  
# Linear Discriminant Analysis: Sad  
```{r lda.sad, echo=FALSE,warning=FALSE,message=FALSE}  
# Data Reshape
d.sad <- data.frame(cbind(cases$class,cases$age,cases$music.yrs,cases$gender,acu.sad)) 
colnames(d.sad) <- c("type","age","music.yrs","gender","nb0","nb2","nb4","nb8","nb16","nb32","nb64")
d.sad$type <- cases$class
d.sad$id <- rownames(d.sad)

# Linear Discriminant model MASS library
m.sad <- lda(type~nb0+nb2+nb4+nb8+nb16+nb32+nb64, d.sad)
# Obtains the predictor from the LDA to an object
p.sad <- predict(m.sad)
# Accuracy of prediction
freqtable <- table(p.sad$class, d.sad$type)
# Percent correct of each category of G
pander(diag(prop.table(freqtable)),"Accuracy prediction: Sad")
pander("  \n  ")

# Frecuency table
rownames(freqtable) <- paste0("Predicted ", d.sad$type %>% levels)
freqtable %>% addmargins %>% pander("LDA sad: Observed vs. Predicted Frequencies")
pander("  \n  ")

# Total percent correct
pander(paste("Sad total % correct:",sum(diag(prop.table(freqtable)))))
pander("  \n  ")

# Proportions
prop.table(freqtable) %>% addmargins %>% pander("LDA sad: Observed vs. Predicted Proportions %")
pander("  \n  ")

# LDA to variables
d.sad$LDA1 <- p.sad$x[,1]
d.sad$LDA2 <- p.sad$x[,2]
d.sad$LDA3 <- p.sad$x[,3]

# LDA table 
gMeans <- d.sad %>% group_by(type) %>% select(LDA1,LDA2,LDA3) %>%  summarise_each(funs(mean)) 
gMeans %>% pander("Happy: group means by LDA")
pander("  \n  ")

# PLot the LDA
grid.arrange(
ggplot(d.sad, aes(LDA1, fill = type)) + geom_density(alpha = 0.5, color = NA) + theme(legend.position = "top"),
ggplot(d.sad, aes(LDA2, fill = type)) + geom_density(alpha = 0.5, color = NA) + theme(legend.position = "top"),
ggplot(d.sad, aes(LDA3, fill = type)) + geom_density(alpha = 0.5, color = NA) + theme(legend.position = "top"),
# Plot the LDA1 vs LDA2 with color per group
ggplot(d.sad, aes(LDA1, LDA2, color = type)) + 
  geom_point(alpha = 0.5) + 
  geom_text(data = gMeans, 
            aes(label = type),
            color = "black", 
            vjust = 1.75) + 
  geom_point(data = gMeans, 
             aes(fill = type), 
             size = 4, 
             color = "black", 
             pch = 21) + 
  theme(legend.position = "none") +
  coord_equal(),
ncol=2, nrow = 2)
```
  
  
```{r lda.sad.plots, echo=FALSE,warning=FALSE,message=FALSE, fig.height= 2.5}
# Plots the discriminat functions
par(mfrow=c(1,3))
# scater plot of each subject by LDA value
scatterplot3d(d.sad$LDA1,d.sad$LDA2,d.sad$LDA3,color = c("red","green","blue","orange")[d.happy$type],pch = 19,box = FALSE,main = "LDA Sad", xlab = "LDA1", ylab = "LDA2", zlab = "LDA3")
addgrids3d(d.sad$LDA1,d.sad$LDA2,d.sad$LDA3, grid = c("xy", "xz", "yz"))
# Linear discrimant functions
drawparti(grouping = d.sad$type, x = d.sad$LDA1, y = d.sad$LDA2, xlab = "LDA1", ylab = "LDA2")
drawparti(grouping = d.sad$type, x = d.sad$LDA2, y = d.sad$LDA3, xlab = "LDA2", ylab = "LDA3")
```

\clearpage

# Figure 4. Cannonical Discriminant Analysis (CDA)
```{r CDA, echo=FALSE,warning=FALSE,message=FALSE}  
# Linear models for each emotion
lm.sad   <- lm(cbind(nb0, nb2, nb4, nb8, nb16, nb32, nb64) ~ type, d.sad)
lm.happy <- lm(cbind(nb0, nb2, nb4, nb8, nb16, nb32, nb64) ~ type, d.happy)

# Canonical discriminant analysis
cda.happy <- lm.happy %>% candisc 
cda.sad <- lm.sad %>% candisc 
```

Two generalized canonical discriminant analysis was compute using the multivariate linear model:  
$$Group~nb0+nb2+nb4+nb8+nb16+nb32+nb64$$  
to obtain the canonical scores and vectors, one for HAPPY and the other for SAD. It represents a transformation of the original variables in the scpace of maximal differences for the group. 
THe biplot shows the canonical scores for the groups defined by the term as points and the canonical structure coefficients as vectors from the origin.  
  
Standardized beta coefficients are given for each variable in each discriminant (canonical) function, and the larger the standardized coefficient, the greater is the contribution of the respective variable to the discrimination between groups. However, these coefficients do not tell us between which of the groups the respective functions discriminate.  

**Happy standardized coefficients**  
```{r CDA.happy, echo=FALSE,warning=FALSE,message=FALSE}  
# Standardized coefficients of the CDA
cda.happy$coeffs.std %>% pander("Happy CDA standardized coefficients")
```

  
**Sad standardized coefficients**
```{r CDA.sad, echo=FALSE,warning=FALSE,message=FALSE}  
# Standardized coefficients of the CDA
cda.sad$coeffs.std %>% pander("Sad CDA standardized coefficients")
```

  
The discriminant function coefficients denote the unique contribution of each variable to the discriminant function, while the structure coefficients denote the simple correlations between the variables and the functions

```{r CDA.res, echo=FALSE,warning=FALSE,message=FALSE, fig.height=3.5} 
# Black BackGround Settings
par(bg="black", col.lab="white", col.axis = 'white', col.main= "white", fg="white",mfrow=c(1,2))

# FIGURE 4 - Canonical Discriminant Analysis
myColors <- hsv((c(0,80,160,240) + 80)/360,s = 0.8,v = 0.8,0.7)
plot(cda.happy, col = myColors, pch = rep(16,4), bty = "n", xpd = T,var.col = "red3",ylim=c(-5,5),xlim=c(-6,4), scale=5,main="Happy")
plot(cda.sad, col = myColors, pch = rep(16,4), bty = "n", xpd = T,var.col = "red3",ylim=c(-5,5),xlim=c(-6,4), scale=5,main="Sad")
```


